---
title: "Problem Set 0"
author: "Kristin Gasca"

output:
  pdf_document: default
  html_notebook: default
---




# Problem 1: Drawing Points

In this problem, we will work with 5 points labelled A, B, C, D, and E:

| Point | $X$ | $Y$ | `pch` | `cex` |     `col`    |
|:-----:|:---:|:---:|:-----:|:-----:|:-------------|
|   A   |  1  |  7  |   19  |   3   | "navy"       |
|   B   |  4  |  3  |   15  |   6   | "blueviolet" |
|   C   |  6  |  5  |   17  |   4   | "firebrick2" |
|   D   |  2  |  2  |   18  |   5   | "cyan3"      |
|   E   |  7  |  4  |   16  |   2   | "seagreen3"  |

## Part (a)

Create an empty plot with no data, where the $x$-axis ranges from 0 to 8, and the $y$-axis ranges from 0 to 10. Be sure to specify a main title, as well as titles for the $x$-axis and $y$-axis. Then use the `points()` function 5 times to graph each of the 5 points A, B, C, D, and E individually, using the point shape, size, and color specified in the table for each point. Finally, annotate the points with text; you can use either the labels of the points (A, B, C, etc.) or the coordinates of the points.

**Solution**

```{r}
plot(
    x = NULL,
    xlim = c(0, 8),
    ylim = c(0, 10),
    main = "Plot for Part 1: Part (a)",
    xlab = "x axis",
    ylab = "y axis"
)

# Point A 
points(
    x = 1,
    y = 7, 
    pch = 19,
    cex = 3,
    col = "navy"
)

text(
    x = 1, 
    y = 7.8, 
    labels = "(1, 7)"
)


# Point B
points(
    x = 4,
    y = 3, 
    pch = 15,
    cex = 6,
    col = "blueviolet"
)

text(
    x = 4, 
    y = 4.3, 
    labels = "(4, 3)",
)


# Point C
points(
    x = 6,
    y = 5, 
    pch = 17,
    cex = 4,
    col = "firebrick2"
)

text(
    x = 6, 
    y = 6.3, 
    labels = "(6, 5)",
)


# Point D 
points(
    x = 2,
    y = 2, 
    pch = 18,
    cex = 5,
    col = "cyan3"
)

text(
    x = 2, 
    y = 3, 
    labels = "(2, 2)",
)


# Point E
points(
    x = 7,
    y = 4, 
    pch = 16,
    cex = 2,
    col = "seagreen3"
)

text(
    x = 7, 
    y = 4.6, 
    labels = "(7, 4)",
)
```




## Part (b)

Create an empty plot with no data, where the $x$-axis ranges from 0 to 8, and the $y$-axis ranges from 0 to 10. Be sure to specify a main title, as well as titles for the $x$-axis and $y$-axis. Then use the `points()` function exactly once to graph the 5 points A, B, C, D, and E. (Hint: you'll have to use vectors.) Your graph should look exactly the same as the one you created in part (a). Finally, annotate the points with text; you can use either the labels of the points (A, B, C, etc.) or the coordinates of the points.

**Solution**

```{r}
# Create an empty plot with 5 points using VECTORS 
plot(
    x = NULL,
    xlim = c(0, 8),
    ylim = c(0, 10),
    main = "Plot for Part 1: Part (b)",
    xlab = "x axis",
    ylab = "y axis"
)

points(
  x = c(1, 4, 6, 2, 7),
  y = c(7, 3, 5, 2, 4),
  pch = c(19, 15, 17, 18, 16), 
  cex = c(3, 6, 4, 5, 2),
  col = c(
    'navy', 
    'blueviolet',
    'firebrick2',
    'cyan3',
    'seagreen3'
  )
)

# Add same labels, using same exact coordinates from part a
text(
  x = c(1, 4, 6, 2, 7),
  y = c(7.8, 4.3, 6.3, 3, 4.6),
  labels = c('(1, 7)', '(4, 3)', '(6, 5)', '(2, 2)', '(7, 4)'),
)
```




## Part (c)

Use a single call to the `plot()` function to create the plotting region and plot all 5 points. As before, the $x$-axis ranges from 0 to 8, and the $y$-axis ranges from 0 to 10. Be sure to specify a main title, as well as titles for the $x$-axis and $y$-axis. Your graph should look the same as the graphs in parts (a) and (b).

**Solution**

```{r}
plot(
    # use vector for all points instead of creating an empty plot
    x = c(1, 4, 6, 2, 7),
    y = c(7, 3, 5, 2, 4),
    # x and y ranges remain the same
    xlim = c(0, 8),
    ylim = c(0, 10),
    # same labels remain
    main = "Plot for Part 1: Part (c)",
    xlab = "x",
    ylab = "y",
    # add styling options
    pch = c(19, 15, 17, 18, 16), 
    cex = c(3, 6, 4, 5, 2),
    col = c(
      'navy', 
      'blueviolet',
      'firebrick2',
      'cyan3',
      'seagreen3'
    ),
)

# Add same labels, using same exact coordinates from part a and b
text(
  x = c(1, 4, 6, 2, 7),
  y = c(7.8, 4.3, 6.3, 3, 4.6),
  labels = c('(1, 7)', '(4, 3)', '(6, 5)', '(2, 2)', '(7, 4)'),
)
```


\newpage

End of problem 1

\newpage




# Problem 2: TIME

Create an empty plot with no data. You'll have to decide on the ranges of the $x$- and $y$-axes. Use "TIME" for the main title, but just use empty strings for the $x$- and $y$-axis titles. Then, using only `segment()` commands, draw the letters of the word "TIME". By my count, you'll need at least 13 separate `segment() commands, although you are welcome and encouraged to use more. Try to do something creative with the line type, width, and color. You'll get full credit as long as we can see the letters for the word "TIME", but you'll get more out of the problem if you take an extra 10 minutes to explore this a little bit.

**Solution**

```{r}
plot(
    # Create an empty plot with no data, add a main title, and select ranges
    x = NULL,
    xlim = c(0, 30),
    ylim = c(0, 8),
    main = "TIME",
    axes = FALSE,
    # Use empty strings for the x and y axis titles
    xlab = " ",
    ylab = " ", 
)

# T SEGMENTS
segments(
  x0 = 0,
  y0 = 8,
  x1 = 4,
  y1 = 8, 
  # style options
  lwd = 9,
  col = 'red',
)

segments(
  x0 = 2,
  y0 = 7.5,
  x1 = 2,
  y1 = 0, 
  # style options
  lwd = 9,
  col = 'orange'
)

# I SEGMENTS
segments(
  x0 = 5,
  y0 = 8,
  x1 = 9,
  y1 = 8, 
  # style options
  lwd = 9,
  col = 'seagreen3'
)

segments(
  x0 = 7,
  y0 = 7.5,
  x1 = 7,
  y1 = 0.5, 
  # style options
  lwd = 9,
  col = 'darkgreen'
)

segments(
  x0 = 5,
  y0 = 0,
  x1 = 9,
  y1 = 0, 
  # style options
  lwd = 9,
  col = 'seagreen3'
)

# M SEGMENTS
segments(
  x0 = 10,
  y0 = 0,
  x1 = 12,
  y1 = 8, 
  # style options
  lwd = 9,
  col = 'dodgerblue3'
)

segments(
  x0 = 15,
  y0 = 0,
  x1 = 12.5,
  y1 = 8, 
  # style options
  lwd = 9,
  col = 'purple'
)

segments(
  x0 = 15.5,
  y0 = 0,
  x1 = 18.5,
  y1 = 8, 
  # style options
  lwd = 9,
  col = 'lavender'
)

segments(
  x0 = 21,
  y0 = 0,
  x1 = 19,
  y1 = 8, 
  # style options
  lwd = 9,
  col = 'lightpink'
)

# E SEGMENTS
segments(
  x0 = 23,
  y0 = 0,
  x1 = 23,
  y1 = 8, 
  # style options
  lwd = 9,
  col = 'hotpink'
)

segments(
  x0 = 23.5,
  y0 = 8,
  x1 = 27.5,
  y1 = 8, 
  # style options
  lwd = 9,
  col = 'gray'
)

segments(
  x0 = 23.5,
  y0 = 4,
  x1 = 27.5,
  y1 = 4, 
  # style options
  lwd = 9,
  col = 'lightblue'
)

segments(
  x0 = 23.5,
  y0 = 0,
  x1 = 27.5,
  y1 = 0, 
  # style options
  lwd = 9,
  col = 'gold'
)
```






\newpage

End of problem 2

\newpage







# Problem 3: Triangles

In this problem, we'll draw the same triangle using three different techniques. The triangle will have vertices at three points denoted $A$, $B$, and $C$, with these coordinates:

| Point | $x$ | $y$ |
|:-----:|:---:|:---:|
| A | 1 | 2 |
| B | 2 | 7 |
| C | 5 | 3 |


## Part (a): Using line segments

In the first method, we'll draw a triangle by using the `segments()` function three times:

* First, create an empty plot with no data, where the $x$-axis ranges from 0 to 6 and the $y$-axis ranges from 0 to 8:

* Second, draw a line segment from the point $A = (1, 2)$ to the point $B = (2, 7)$.

* Next, draw a line segment from the point $B = (2, 7)$ to the point $C = (5, 3)$.

* Finally draw a line segment from the point $C = (5, 3)$ back to $A = (1, 2)$.

The lines of the triangle should be colored blue, with a line width of 2. Finally, draw a black circular point at each of the vertices of the triangle.

**Solution**

```{r}
plot(
    # Create an empty plot with no data, add a main title, and specified ranges
    x = NULL,
    xlim = c(0, 6),
    ylim = c(0, 8),
    main = "Triangles Using Segments",
    xlab = "x",
    ylab = "y", 
    type = 'b'
)

# Draw a line segment from point A to B, B to C, and C to A
segments(
  x0 = 1, 
  y0 = 2,
  x1 = 2, 
  y1 = 7,
  col = 'blue',
  lwd = 2, 
)

segments(
  x0 = 2, 
  y0 = 7,
  x1 = 5, 
  y1 = 3,
  col = 'blue',
  lwd = 2, 
)

segments(
  x0 = 5, 
  y0 = 3,
  x1 = 1, 
  y1 = 2,
  col = 'blue',
  lwd = 2, 
)

points(
  x = c(1, 2, 5),
  y = c(2, 7, 3),
  pch = 19,
  cex = 2,
  col = 'black', 
)
```


## Part (b): Using the `lines()` function

Now we'll draw the same triangle again, with the same points for the vertices, but this time we'll use the `lines()` function. Remember that this function always draws *connected* line segments. Remember to connect the last point $C$ to the first point $A$! The lines of the triangle should be colored blue with a line width of 2. As in part (a), you should also draw points at the vertices; here you are welcome to copy and paste your code from part (a). Start with the same empty plot with no data that you used in part (a).

**Solution**

```{r}
plot(
    # Create an empty plot with no data, add a main title, and specified ranges
    x = NULL,
    xlim = c(0, 6),
    ylim = c(0, 8),
    main = "Triangles Using the lines() Function",
    xlab = "x",
    ylab = "y"
)

lines(
  x = c(1, 2, 5, 1),
  y = c(2, 7, 3, 2),
  col = 'blue',
  lwd = 2, 
)

points(
  x = c(1, 2, 5),
  y = c(2, 7, 3),
  pch = 19,
  cex = 2,
  col = 'black', 
)
```




## Part (c): Using polygons

Finally, draw the triangle using the `polygon()` function. Remember with this function that it draws connected line segments, and that the last point $C$ is automatically connected to the first point $A$. The lines of the triangle should be colored blue with a line width of 2. Decorate your triangle with points just like in parts (a) and (b).

**Solution**

```{r}
plot(
    # Create an empty plot with no data, add a main title, and specified ranges
    x = NULL,
    xlim = c(0, 6),
    ylim = c(0, 8),
    main = "Triangles Using the lines() Function",
    xlab = "x",
    ylab = "y"
)

polygon(
  x = c(1, 2, 5),
  y = c(2, 7, 3),
  border = 'blue',
  lwd = 2, 
)

points(
  x = c(1, 2, 5),
  y = c(2, 7, 3),
  pch = 19,
  cex = 2,
  col = 'black', 
)
```


\newpage

End of Problem 3

\newpage




# Problem 4: Sales Chart

A grocery store has data on the sales of three breakfast cereals: Sugar Bomz (SBZ), Krispee Yummm (KYM), and Healthy Kale and Tofu (HKT):

| Year |  SBZ  |  KYM  |  HKT  |
|:----:|:-----:|:-----:|:-----:|
| 2016 |  40.3 |  33.7 |  20.4 |
| 2017 |  55.2 |  38.9 |  15.8 |
| 2018 |  63.4 |  41.4 |  10.6 |
| 2019 |  76.3 |  44.6 |  11.5 |

Create a sales chart for this data. The $x$-axis should indicate the year, and the $y$-axis should indicate the sales volume. Draw the sales for each brand as a connected line, and use a different point shape and color for each brand. Remember to include a main title, axis titles, and a legend.

**Solution**

```{r}
plot(
  x = NULL,
  xlim = c(2015, 2020),
  ylim = c(5, 80),
  xlab = 'Year',
  ylab = 'Sales',
  main = "Cereal Sales for Three Brands (2016-2019)",
  las = 1
)

# Line segment and points for SBZ Cereal
lines(
  x = c(2016, 2017, 2018, 2019),
  y = c(40.3, 55.2, 63.4, 76.3),
  col = 'lavender',
  lwd = 3,
  lty = 'dashed'
)

points(
  x = c(2016, 2017, 2018, 2019),
  y = c(40.3, 55.2, 63.4, 76.3),
  col = 'lavender',
  cex = 3,
  pch = 19,
)

# Line segment and points for KYM Cereal
lines(
  x = c(2016, 2017, 2018, 2019),
  y = c(33.7, 38.9, 41.4, 44.6),
  col = 'hotpink',
  lwd = 3,
  lty = 'dotted'
)

points(
  x = c(2016, 2017, 2018, 2019),
  y = c(33.7, 38.9, 41.4, 44.6),
  col = 'hotpink',
  cex = 2,
  pch = 15,
)

# Line segment and points for HKT Cereal
lines(
  x = c(2016, 2017, 2018, 2019),
  y = c(20.4, 15.8, 10.6, 11.5),
  col = 'lightblue',
  lwd = 3,
  lty = 'solid', 
)

points(
  x = c(2016, 2017, 2018, 2019),
  y = c(20.4, 15.8, 10.6, 11.5),
  col = 'lightblue',
  pch = 17,
  cex = 3
)


# Create a legend
legend(
  x = 2015,
  y = 80,
  legend = 
    c(
      'SBZ',
      'KYM',
      'HKT'
    ),
  lwd = 3,
  lty = 
    c(
      'dashed', 
      'dotted',
      'solid'
    ),
  col = c('lavender', 'hotpink', 'lightblue'), 
  pch = c(19, 15, 17)
)
```



\newpage

End of problem 4

\newpage




# Problem 5: Graphing Functions

In this problem, we will gain experience graphing polynomial functions.

## Part (a): Quadratic Functions

First, let's graph this function:
$$
y\ =\ 2 x^2 - 2 x - 4
$$
Make your graph using these steps:

* First, create an empty plot with $x$ ranging from -3 to +3 and $y$ ranging from -5 to +5. Label the $x$ axis with "x" and the $y$ axis with "f(x)".

* Draw in the horizontal reference line $y = 0$ from $x = -3$ to $x = +3$.

* Draw in the vertical reference line $x = 0$ from $y = -5$ to $y = +5$.

* Draw the graph of the function using the `curve()` command. (Hint: don't forget to use the `add = TRUE` option!) The function curve should be a solid line of width 2, and should be colored blue.

* Draw points at the locations $(-1, 0)$ and $(2, 0)$, which represent the roots of the quadratic function.

**Solution**

```{r}
# Create an empty plot with the specified ranges and labels
plot(
  x = NULL,
  xlim = c(-3, 3),
  ylim = c(-5, 5),
  xlab = "x", 
  ylab = "f(x)"
)

# Horizontal reference line
segments(
  x0 = -3, 
  y0 = 0,
  x1 = 3, 
  y1 = 0,
  lty = 'solid',
  lwd = 2,
  col = 'gray50'
)

# Vertical reference line
segments(
  x0 = 0, 
  y0 = -5,
  x1 = 0, 
  y1 = 5,
  lty = 'solid',
  lwd = 2,
  col = 'gray50'
)

# Graph f(x) = 
curve(
  expr = (2*x^2) - (2*x) - (4),
  lwd = 2, 
  lty = 'solid',
  col = 'blue',
  add = TRUE
)

points(
  x = c(-1, 2),
  y = c(0, 0),
  col = 'blue',
  cex = 2
)
```




## Part (b): Cubic functions

Now let's draw the graph of a cubic equation:
$$
y\ =\ 2 x^3 - 2 x
$$
Make your graph using these steps:

* First, create an empty plot with $x$ ranging from -3 to +3 and $y$ ranging from -5 to +5. Label the $x$ axis with "x" and the $y$ axis with "f(x)".

* Draw in horizontal and vertical reference lines.

* Using a solid blue line of width 3, draw the graph of the function using the `curve()` command.

* Draw points at the locations (-1, 0), (0, 0), and (1, 0), which represent the roots of the cubic function.

**Solution**

```{r}
# Create an empty plot with the specified ranges and labels
plot(
  x = NULL,
  xlim = c(-3, 3),
  ylim = c(-5, 5),
  xlab = "x", 
  ylab = "f(x)"
)

# Horizontal reference line
segments(
  x0 = -3, 
  y0 = 0,
  x1 = 3, 
  y1 = 0,
  lty = 'solid',
  lwd = 2,
  col = 'gray50'
)

# Vertical reference line
segments(
  x0 = 0, 
  y0 = -5,
  x1 = 0, 
  y1 = 5,
  lty = 'solid',
  lwd = 2,
  col = 'gray50'
)

# Graph 
curve(
  expr = (2*x^3) - (2*x),
  lwd = 3, 
  lty = 'solid',
  col = 'blue',
  add = TRUE
)

points(
  x = c(-1, 0, 1),
  y = c(0, 0, 0),
  col = 'blue',
  cex = 2
)
```




## Part (c): Exponential Functions

Create an empty plotting region with no data, where the $x$-axis ranges from 0 to 4 and the $y$-axis ranges from 0 to 2. Include horizontal and vertical reference lines. Also, draw in the dashed horizontal line $y = 1$. Then plot the curves for these functions:

* The function $f(x) = 1 - e^{-x}$.

* The function $g(x) = 1 - e^{-2x}$.

* The function $h(x) = 1 - e^{-3x}$.

Finally, include a main title, axis titles, and a legend.

**Solution**

```{r}
# Create an empty plot with the specified ranges and labels
plot(
  x = NULL,
  xlim = c(0, 4),
  ylim = c(0, 2),
  xlab = "x", 
  ylab = "f(x)", 
  main = "Exponential Functions"
)

# Horizontal reference line
segments(
  x0 = -1, 
  y0 = 0,
  x1 = 5, 
  y1 = 0,
  lty = 'solid',
  lwd = 2,
  col = 'gray50'
)

# Vertical reference line
segments(
  x0 = 0, 
  y0 = -5,
  x1 = 0, 
  y1 = 5,
  lty = 'solid',
  lwd = 2,
  col = 'gray50'
)

# Dashed horizontal line at y = 1 
segments(
  x0 = -1, 
  y0 = 1,
  x1 = 4.5, 
  y1 = 1,
  lty = 'dashed',
  lwd = 2,
  col = 'gray50'
)

# Graphs
curve(
  expr = 1 - exp(-x),
  lwd = 3, 
  lty = 'solid',
  col = 'pink',
  add = TRUE
)

curve(
  expr = 1 - exp(-2 * x),
  lwd = 3, 
  lty = 'solid',
  col = 'blue',
  add = TRUE
)

curve(
  expr = 1 - exp(-3 * x),
  lwd = 3, 
  lty = 'solid',
  col = 'purple',
  add = TRUE
)

# Legend
legend(
  x = 2.5,
  y = 2,
  legend = 
    c(
      'f(x) = 1 - e^{-x}',
      'g(x) = 1 - e^{-2x}',
      'h(x) = 1 - e^{-3x}'
    ),
  lty = 'solid',
  lwd = 3,
  col = c('pink', 'blue', 'purple')
)
```



\newpage

End of problem 5

\newpage




# Problem 6: Normal Distributions

Now we'll explore normal density curves.

## Part (a): A normal distribution

Create an empty plot where the $x$-axis ranges from -2 to 7 and the $y$-axis ranges from 0 to 0.5. Then use a solid red line with a weight of 2 to graph the density curve for a normal distribution with a mean of 2.5 and a standard deviation of 1.

**Solution**

```{r}
plot(
  x = NULL,
  xlim = c(-2, 7),
  ylim = c(0, 0.5),
  xlab = 'x',
  ylab = 'y',
  main = "Normal Distribution Curve",
  las = 1
)

curve(
  expr = dnorm(x, mean = 2.5, sd = 1),
  lwd = 2, 
  col = 'red', 
  add = TRUE
)
```




## Part (b): Another normal distribution

Create an empty plot where the $x$-axis ranges from 2 to 6 and the $y$-axis ranges from 0 to 0.9. Then use a dashed blue line with a weight of 3 to graph the density curve for a normal distribution with a mean of 4 and a standard deviation of 0.5.

**Solution**

```{r}
plot(
  x = NULL,
  xlim = c(2, 6),
  ylim = c(0, 0.9),
  xlab = 'x',
  ylab = 'y',
  main = "Normal Distribution Curve",
  las = 1
)

curve(
  expr = dnorm(x, mean = 4, sd = 0.5),
  lwd = 3, 
  col = 'blue', 
  lty = 'dashed',
  add = TRUE
)
```




## Part (c): Another normal distribution

Create an empty plot where the $x$-axis ranges from 2 to 12 and the $y$-axis ranges from 0 to 0.25. Then use a dotted black with a weight of 4 to graph the density curve of a normal distribution with a mean of 7 and a standard deviation of 2.

**Solution**

```{r}
plot(
  x = NULL,
  xlim = c(2, 12),
  ylim = c(0, 0.25),
  xlab = 'x',
  ylab = 'y',
  main = "Normal Distribution Curve",
  las = 1
)

curve(
  expr = dnorm(x, mean = 7, sd = 2),
  lwd = 4, 
  lty = 'dotted',
  add = TRUE
)
```





## Part (d): Putting it all together

Now create a single graph that has all three curves from parts (a), (b), and (c). Use the same line width, line type, and color settings that you used in those parts. The challenge here is to figure out what the appropriate settings for `xlim` and `ylim` should be so that all three curves are displayed. Remember that you have to choose the ranges of the $x$-axis and the $y$-axis when you create the empty plot, so once you've made that decision you won't be able to change it when drawing the individual curves. (Of course, if you decide that you don't like your choices for the ranges, you can just change them and then re-run the code chunk.)

**Solution**

```{r}
plot(
  x = NULL,
  xlim = c(-2, 12),
  ylim = c(0, 0.9),
  xlab = 'x',
  ylab = 'y',
  main = "Normal Distribution Curve",
  las = 1
)

curve(
  expr = dnorm(x, mean = 2.5, sd = 1),
  lwd = 2, 
  col = 'red', 
  add = TRUE
)

curve(
  expr = dnorm(x, mean = 4, sd = 0.5),
  lwd = 3, 
  col = 'blue', 
  lty = 'dashed',
  add = TRUE
)

curve(
  expr = dnorm(x, mean = 7, sd = 2),
  lwd = 4, 
  lty = 'dotted',
  add = TRUE
)
```




\newpage

End of Problem 6

\newpage





# Problem 7: Mondrian (Extra Credit 0.5 Points)

Time to express our inner souls with some more modern art!

Go to the Wikipedia article on Piet Mondrian, the great Dutch abstract artist, and scroll down to find the image of *Composition II in Red, Blue, and Yellow*, painted in 1930. Your goal is to recreate this painting using the tools that we developed in lecture.

* First, create a completely blank square plot with no axes, with $x$ ranging from 0 to 10 and $y$ ranging from 0 to 10.

* Draw a frame around the plot area by drawing a polygon with vertices at (0, 0), (0, 10), (10, 10), and (10, 0).

* Draw the red, blue, and yellow rectangles using the `polygon()` function. You'll have to figure out the appropriate coordinates!

* Finally, draw in the black lines, using the `lwd =` option to control the width of the lines.

You'll find it useful when drawing thick lines to include this line code at the beginning of the code chunk:

Your picture doesn't have to be perfect, but it should be a reasonable approximation to the original.

**Solution**

```{r}
par( lend = 1, pty = "s")

# Create a completely blank square with no axes and specified range
plot(
    x = NULL,
    xlim = c(0, 10),
    ylim = c(0, 10), 
    axes = FALSE,
    
    # Erase axis titles by assigning empty strings to the x and y lab parameters
    xlab = "",
    ylab = ""
)

#Left hand column graphics
polygon(
  x = c(0, 2, 2, 0),
  y = c(7, 7, 10, 10),
  col = "ivory",
  border = 'black',
  lwd = 4
)

polygon(
  x = c(0, 2, 2, 0),
  y = c(2.5, 2.5, 6.9, 6.9),
  col = "ivory",
  border = 'black',
  lwd = 4
)

polygon(
  x = c(0, 2, 2, 0),
  y = c(0, 0, 2.4, 2.4),
  col = "blue",
  border = 'black',
  lwd = 4
)

# Center graphic 
polygon(
  x = c(2.1, 10, 10, 2.1),
  y = c(2.5, 2.5, 10, 10),
  col = "red",
  border = 'black',
  lwd = 4
)

# Bottom graphics
polygon(
  x = c(2.1, 9, 9, 2.1),
  y = c(0, 0, 2.4, 2.4),
  col = "ivory",
  border = 'black',
  lwd = 4
)

polygon(
  x = c(9.1, 10, 10, 9.1),
  y = c(1.3, 1.3, 2.4, 2.4),
  col = "ivory",
  border = 'black',
  lwd = 4
)

polygon(
  x = c(9.1, 10, 10, 9.1),
  y = c(0, 0, 1.3, 1.3),
  col = "yellow",
  border = 'black',
  lwd = 4
)

# Draw a frame over the polygons by drawing a polygon 
polygon(
  x = c(0, 0, 10, 10),
  y = c(0, 10, 10, 0),
  lwd = 5
)
```











