---
title: "Problem Set 9"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---



Let's clear the global computing environment:
```{r}
rm( list = ls() )
```




# Problem 1: Stock Data

## Part (a): Reading in the stock data

First, read in the data from the file "Problem 1 Stock Data.csv", which is contained in the folder "Problem Set 9 Data".

**Solution**

```{r}
stock.data.frame <- read.csv(
    "../Problem Set 9 Data/Problem 1/Problem 1 Stock Data.csv"
)
```



## Part (b): Defining the VWAP function

In Problem Set 7, Problem 5, part (a), you defined a function to calculate the volume-weighted average price (VWAP), given a vector of price per share data and a vector of number of shares sold data. Copy that function, and paste it into this file; if you want to, you can use my solution, but you'll get more out of the problem if you use your own code.

**Solution**

```{r}
vwap <- function(
  price.per.share,
  number.of.shares.sold) {
  return(
      sum(price.per.share * number.of.shares.sold) /
      sum(number.of.shares.sold)
  )
}
```



## Part (c): Selecting rows

Using the data frame from part (a), select the rows corresponding to sales for Global Widget. Save this data frame in a variable, and display the first 6 rows using a `head()` statement.

**Solution**

```{r}
global.widget.data.frame <- 
    stock.data.frame[stock.data.frame$company.name == "Global Widget",]

head(global.widget.data.frame)
```



## Part (d): Selecting columns

Using the Global Widget data frame you created in part (c), select the columns for price per share data and number of shares sold, and use your function from part (b) to calculate the VWAP for Global Widget. Report your result using a `cat()` statement, displaying this value to 2 decimal places.

**Solution**

```{r}
vwap.global.widget <- vwap(
    price.per.share = global.widget.data.frame$price.per.share,
    number.of.shares.sold = global.widget.data.frame$number.of.shares.sold
)

cat(
    "Global Widget VWAP:",
    formatC(
        vwap.global.widget,
        format = 'f',
        digits = 2
    )
)
```



## Part (e): Writing a function

Write a function that takes a character string identifier for a company name, and returns the VWAP for the company:

* First, using the stock data frame from part (a), the function selects all transactions for the specified company.

* Next, the function computes the VWAP for this company using the function definition form part (b) and the appropriate rows from the data frame, and then returns this value.

There's nothing to report for this part, but write your code clearly so the TAs can understand what you're doing.

**Solution**

```{r}
vwap.calc.function <- function(company.name) {
    company.stock <- stock.data.frame[
        stock.data.frame$company.name == company.name,
        ]
    return(
        vwap(
            price.per.share = company.stock$price.per.share,
            
            number.of.shares.sold = company.stock$number.of.shares.sold
        )
    )
}
```



## Part (f): WiDgT VWAP

Use your vwap database calculator function to determine the VWAP for WiDgT. Report this value using a `cat()` statement, displaying this value with 2 decimal places.

**Solution**

```{r}
widgt.vwap <- vwap.calc.function("WiDgT")

cat(
    "WiDgT vwap:",
    formatC(
        widgt.vwap,
        format = 'f',
        digits = 2
    )
)
```



\newpage

End of problem 1

\newpage





# Problem 2: Baseball Database Reporter

In this problem, we're going to pull together a number of ideas from the course to make a baseball database reporter function. The actual new code that you have to write is not all that much.

Our goal in this problem is to create a baseball reporter system that can take the name of a player, query a database for the annual batting data for that player, and then produce a summary report.

We'll start by working through a single case, and then we'll write a function to generalize this method.


## Part (a): Read in the data

First, read in the data from the file "Baseball Batting Database.csv", which is contained in the folder "Problem Set 9 Data".

**Solution**

```{r}
baseball.data.frame <- read.csv(
    "../Problem Set 9 Data/Problem 2/Baseball Batting Database.csv"
)
```



## Part (b): Selecting the rows for Willie Mays

We're going to start by learning how to handle a special case, which for this problem will be Willie Mays.

Select the rows from the baseball database corresponding to Willie Mays. Save this data frame in a variable, and display the first 5 rows directly.

**Solution**

```{r}
willie.mays.data <- baseball.data.frame[baseball.data.frame$Name == "Willie Mays",]
```



## Part (c): Willie Mays baseball report

In Problem Set 7, we constructed a baseball reporter function that took baseball batting data for a player and generated a batting summary report.

Copy the code for this baseball reporter function from your Problem Set 7 solutions, along with any additional functions that you need. (If you want to, you can use the code from the Problem Set 7 solutions.)

Then call the baseball reporter function using the data from part (b) to generate a baseball batting report on Willie Mays.

**Solution**

```{r}
batting.average <-
    function( at.bats, hits ) {
        return(
            hits / at.bats
        )
    }

on.base.percentage <-
    function(
        at.bats,
        hits,
        bob,
        hbp,
        sf
    ) {
        return(
            (hits + bob + hbp) / (at.bats + bob + hbp + sf)
        )
    }

slugging.percentage <-
    function(
        at.bats,
        hits,
        doubles,
        triples,
        home.runs
    ) {
        singles <- hits - (doubles + triples + home.runs)
        
        total.bases <-
            (1 * singles) +
            (2 * doubles) +
            (3 * triples) +
            (4 * home.runs)
        
        return(
            total.bases / at.bats
        )
    }
```

```{r}
# Write the baseball reporter function 
baseball.reporter.function <-
    function(
        name,
        at.bats,
        hits,
        doubles,
        triples,
        home.runs,
        bob,
        hbp,
        sf
    ) {
        outs <- (at.bats - hits)
        
        singles <- (hits - (doubles + triples + home.runs))
        
        player.batting.average <-
            batting.average(
                at.bats = at.bats,
                hits = hits
            )
        
        player.on.base.percentage <-
            on.base.percentage(
                at.bats = at.bats,
                hits = hits,
                bob = bob,
                hbp = hbp,
                sf = sf
            )
        
        player.slugging.percentage <-
            slugging.percentage(
                at.bats = at.bats,
                hits = hits,
                doubles = doubles,
                triples = triples,
                home.runs = home.runs
            )
      
        cat("Player:", name)
        
        cat("\n\n\tAt-bats:", at.bats)
        
        cat("\n\tOuts:", outs)
        
        cat("\n\tBases on balls:", bob)
        
        cat("\n\tHits:", hits)
        
        cat("\n\tSingles:", singles)
        
        cat("\n\tDoubles:", doubles)
        
        cat("\n\tTriples:", triples)
        
        cat("\n\tHome runs:", home.runs)
        
        cat(
            "\n\n\tBatting average:",
            formatC(
                    player.batting.average,
                    format = "f",
                    digits = 3
                )
        )
        
        cat(
            "\n\tOn-base percentage:",
            substr(
                formatC(
                    player.on.base.percentage,
                    format = "f",
                    digits = 3
                ),
            start = 2,
            stop = 5
            )
        )
        
        cat(
            "\n\tSlugging percentage:",
            substr(
                formatC(
                    player.slugging.percentage,
                    format = "f",
                    digits = 3
                ),
            start = 2,
            stop = 5
            )
        )
    }
```

```{r}
baseball.reporter.function(
    name = "Willie Mays",
    at.bats = sum( willie.mays.data$AB ),
    hits = sum( willie.mays.data$H ),
    doubles = sum( willie.mays.data$X2B ),
    triples = sum( willie.mays.data$X3B ),
    home.runs = sum( willie.mays.data$HR ),
    bob = sum( willie.mays.data$BB ),
    hbp = sum( willie.mays.data$HBP ),
    sf = sum( willie.mays.data$SF )
)
```



## Part (e): Player reporter function

Write a function that takes a player's name, and then generates a baseball report:

* First, the function selects the rows from the baseball database corresponding to that player.

* Then the function calls the baseball reporter function, using the data from the selected rows.

There's nothing to report here, but write your code clearly so the TAs can understand what you're doing.

**Solution**

```{r}
player.reporter.function <- function(name) {
    player.data.frame <- 
        baseball.data.frame[baseball.data.frame$Name == name,]

    baseball.reporter.function(
        name = name, 
        at.bats = sum(player.data.frame$AB),
        hits = sum(player.data.frame$H),
        doubles = sum(player.data.frame$X2B),
        triples = sum(player.data.frame$X3B),
        home.runs = sum(player.data.frame$HR),
        bob = sum(player.data.frame$BB),
        hbp = sum(player.data.frame$HBP),
        sf = sum(player.data.frame$SF)
    )
}
```



## Part (f): Roberto Clemente batting report

Use the baseball player reporter function you defined in part (e) to generate a report for Roberto Clemente.

**Solution**

```{r}
player.reporter.function("Roberto Clemente")
```



\newpage

End of problem 2

\newpage



# Problem 3: Sales Data

Sales data for the months of June, July, and August are contained in these files:

* "Problem 3 June Data.csv"

* "Problem 3 July Data.csv"

* "Problem 3 August Data.csv"

These files are contained in the folder "Problem 3" in the "Problem Set 9 Data" folder.



## Part (a): Reading the June data

Read in the file "Problem 3 June Data.csv" and store the data frame in a variable. Then display the first 8 rows using a `head()` statement.

**Solution**

```{r}
june.data <- read.csv(
    "../Problem Set 9 Data/Problem 3/Problem 3 June Data.csv"
)
```



## Part (b): Reading the July data

Read in the file "Problem 3 July Data.csv" and store the data frame in a variable. Then display the first 8 rows using a `head()` statement.

**Solution**

```{r}
july.data <- read.csv(
    "../Problem Set 9 Data/Problem 3/Problem 3 July Data.csv"
)
```



## Part (c): Reading in the August data

Read in the file "Problem 3 August Data.csv" and store the data frame in a variable. Then display the first 8 rows using a `head()` statement.

**Solution**

```{r}
august.data <- read.csv(
    "../Problem Set 9 Data/Problem 3/Problem 3 August Data.csv"
)
```



## Part (d): Appending rows

Combine these three monthly data sets together by appending rows into a single data frame and store it in a variable. You might have to make some adjustments before you're able to do this. The column names for this combined data frame should be:

* Location

* Revenues

* Costs

Once you've created the combined data frame, display the first 8 rows using a `head()` statement.

**Solution**

```{r}
# Column names
names(june.data) <- c('Location', 'Revenues', 'Costs')

names(july.data) <- c('Location', 'Revenues', 'Costs')

names(august.data) <- c('Location', 'Revenues', 'Costs')

# Combine the 3 data sets
combined.months.data <- rbind(
    june.data,
    july.data,
    august.data
)
```



## Part (e): Calculating profits

Create a new column in the combined data frame from part (d) named "Profits", defined as the Revenues minus the Costs. Then display the first 8 rows using a `head()` statement.

**Solution**

```{r}
combined.months.data$Profits <- 
    (combined.months.data$Revenues - combined.months.data$Costs)
```



## Part (f): Histogram of profits

Using the profit data from part (e), construct a histogram of the profits for the months June, July, and August.

**Solution**

```{r}
#length(combined.months.data$Profits)

hist(
    combined.months.data$Profits,
    main = "Histograms of Profits",
    xlab = 'profits',
    ylab = 'frequency',
    breaks = 50,
    col = 'hotpink'
)
```



\newpage

End of problem 3

\newpage



# Problem 4: Cereal Data

## Part (a): Reading the brand data

Read in the file "Problem 4 Brand Data.csv", which is located in the folder "Problem 4". Save this data frame in a variable, and display the first 8 rows using a `cat()` statement.

**Solution**

```{r}
problem.4.brand.data <- read.csv(
    "../Problem Set 9 Data/Problem 4/Problem 4 Brand Data.csv"
)
```



## Part (b): Reading the sales data

Read in the file "Problem 4 Sales Data.csv", which is located in the folder "Problem 4". Save this data frame in a variable, and display the first 8 rows using a `cat()` statement.

**Solution**

```{r}
problem.4.sales.data <- read.csv(
    "../Problem Set 9 Data/Problem 4/Problem 4 Sales Data.csv"
)
```



## Part (c): Stripping the id numbers

We would like to merge the two data files together, but unfortunately the id numbers for the transactions have extra characters which will prevent this.

Create a new column in the brand data frame which consists of just the transaction id numbers. To do this, use the `substr()` function to strip out the extra characters.

When you're done, display the first 8 rows of the brand data frame, including the column with the transaction id numbers.

**Solution**

```{r}
problem.4.brand.data$transaction.id <- substr(
    problem.4.brand.data$brand.transaction.id,
    start = 1,
    stop = 3
)

head(problem.4.brand.data, n = 8)
```



## Part (d): Stripping the id numbers

We would like to merge the two data files together, but unfortunately the id numbers for the transactions have extra characters which will prevent this.

Create a new column in the sales data frame which consists of just the transaction id numbers. To do this, use the `substr()` function to strip out the extra characters.

When you're done, display the first 8 rows of the sales data frame, including the column with the transaction id numbers.

**Solution**

```{r}
problem.4.sales.data$transaction.id <- substr(
    problem.4.sales.data$sales.transaction.id,
    start = 1,
    stop = 3
)

head(problem.4.sales.data, n = 8)
```



## Part (e): Merging the data frames

Merge the two data frames from parts (c) and (d) together. Save the resulting data frame in a variable, and display the first 8 rows using a `head()` statement.

**Solution**

```{r}
merged.data <- merge(
    problem.4.brand.data,
    problem.4.sales.data,
    by = "transaction.id"
)

head(merged.data, n = 8)
```



## Part (f): Total sales amount

Using your merged data frame, calculate the total sales amount for all the transactions. 

Use the familiar table of cereal brand prices:

| Brand | Price |
|:-----:|:-----:|
|  SBZ  |  2.99 |
|  KYM  |  3.49 |
|  HKT  |  7.99 |

Report your result using a `cat()` statement, rounding to 2 decimal places.

**Solution**

```{r}
lookup.vector <- c(
    'SBZ' = 2.99,
    'KYM' = 3.49,
    'HKT' = 7.99
)

price.per.box <- lookup.vector[merged.data$brand]

total.sales <- sum(price.per.box * merged.data$number.of.boxes.sold)

cat(
    "Total sales: $",
    formatC(
        total.sales,
        format = 'f',
        digits = 2,
        big.mark = ','
    ),
    sep = ''
)
```



\newpage

End of problem 4

\newpage



# Problem 5: Cleaning Baseball Data


To create the datasets of annual baseball batting data, I've used the website `www.baseball-reference.com`, and they actually have an option to convert the data on the website to a dataset in CSV format.


However, there wasn't entirely simple to do, and there were a few issues.


You've already seen one such issue -- sacrifice flies were not recorded until 1954, so this column is reported as `NA`, and this creates issues when we want to calculate the total career number of sacrifice flies.


There were a couple of other issues in the data, but I cleaned them up for you because this was your first time working with datasets, and you didn't have the tools available to you to resolve them.


Now that you've worked through Lecture 10, you have those tools, so we're going to go back and fix these problems.



## Part (a): Ted Williams Batting Data

The file `Ted Williams RAW data.csv` contains the raw batting data that I copied from the `baseball-reference.com` webpage. Load this data, and store it in a variable. Then display the first 5 years of Ted Williams' batting statistics.


**Solution**


```{r}
williams.batting.data <- 
  read.csv(
    "../Problem Set 9 Data/Problem 5/Ted Williams Batting RAW.csv"
  )

# Not including NA years 
head(williams.batting.data[,1:8], 8)
```



## Part (b): Ted Williams at-bat data

Select the data from the column named `AB` as a vector, and store it in a variable. Determine the class of this R object (use the `class()` function from Lecture 1). Then, using complete sentences, write answers for these three questions:

* Since the `AB` column is supposed to be the number of at-bats for each season, what class should this object be?

* What is the actual class of the `AB` vector?

* Why is this happening?

**Solution**

```{r}
at.bat.data <- williams.batting.data$AB

class(at.bat.data)
```

* This column is character based due to the NA values for the years Williams did not play.
* The class should be numeric, as it's a batting statistic. 
* This happens when a single item is a char value, and therefore changes all numeric values to char values. 



## Part (c): Ted Williams' on-base percentage

Using the data frame from part (a), calculate Ted williams' on-base percentage. You'll have to figure out how to fix the problems in part (b), and once you do that you can use the `on.base.percentage()` function included in the CSCI E-5a R Object library. As always, report your result using a `cat()` statement, displaying the value with standard baseball formatting conventions.

**Hint:** Remember the `as.numeric()` function.

**Solution**

```{r}
# Count how many years contain an NA value
# sum(is.na(williams.batting.data))

# Delete those three rows
williams.batting.data <- williams.batting.data[ -(5:7), ]

williams.obp <- on.base.percentage(
    at.bats = sum(as.numeric(williams.batting.data$AB)),
    hits = sum(as.numeric(williams.batting.data$H)),
    bob = sum(as.numeric(williams.batting.data$BB)),
    hbp = sum(as.numeric(williams.batting.data$HBP)),
    sf = sum(as.numeric(williams.batting.data$SF), na.rm = TRUE)
)

cat(
    "Williams' OBP:",
    substr(
        formatC(
            williams.obp,
            format = 'f',
            digits = 3
        ),
        start = 2,
        stop = 5
    )
)
```



## Part (d): Willie Mays' batting data

Your goal in this problem is to calculate Willie Mays' slugging average using the data from `Willie Mays Batting RAW.csv`.

Here's an interesting fact about Willie Mays: he had 10,881 total career at-bats.

If you add up all the numeric values in the column for at-bats, do you get 10,881?

If not, why do you think that is?

Here's another fact about Willie Mays that you might find useful. He began his major league career in 1951 with the New York Giants, who moved to San Francisco in 1958. He played for the San Francisco Giants until he was traded to the New York Mets during the 1972 season, and he finished his playing career with the Mets.

**Solution**

```{r}
mays.batting.data <- read.csv(
    "../Problem Set 9 Data/Problem 5/Willie Mays Batting RAW.csv"
)

mays.at.bats <- mays.batting.data$AB

# Delete NA and TOT rows 
mays.at.bats <- mays.at.bats[ c(-3, -22) ]

sum(as.numeric(mays.at.bats))
```
```{r}
# Calculate Willie Mays' slugging average
slugging.average.function <- function(
    at.bats,
    hits,
    doubles,
    triples,
    home.runs
) {
    singles <- hits - (doubles + triples + home.runs)
    
    total.bases <- 
        (1 * singles) +
        (2 * doubles) +
        (3 * triples) +
        (4 * home.runs)
        
    return(
        total.bases / at.bats
    )
}

# Delete rows 3 and 22 
mays.at.bats <- mays.batting.data$AB[ c(-3, -22) ]
mays.hits <- mays.batting.data$H[ c(-3, -22) ]
mays.doubles <- mays.batting.data$X2B[ c(-3, -22) ]
mays.triples <- mays.batting.data$X3B[ c(-3, -22) ]
mays.home.runs <- mays.batting.data$HR[ c(-3, -22) ]

mays.slugging.average <- slugging.average.function(
    at.bats = sum(as.numeric(mays.at.bats)),
    hits = sum(as.numeric(mays.hits)),
    doubles = sum(as.numeric(mays.doubles)),
    triples = sum(as.numeric(mays.triples)),
    home.runs = sum(as.numeric(mays.home.runs))
)

cat(
    "Willie Mays' slugging average:",
    substr(
        formatC(
            mays.slugging.average,
            format = "f",
            digits = 3
        ),
        start = 2,
        stop = 5
    )
)
```



\newpage

End of Problem 6

\newpage





# Problem 6: Stratified Charts

## Part (a): Stratified boxplot

Construct a stratified boxplot across species for the `Petal.Width` values in the `iris` data frame.

**Solution**

```{r}
boxplot(
    ylim = c(0, 3),
    iris$Petal.Width ~ iris$Species,
    main = "Petal Width Boxplot",
    ylab = 'Width',
    col = c('red', 'yellow', 'blue'),
    las = 1
)
```



## Part (b): Stratified stripchart

Construct a stratified stripchart across species for the `Petal.Width` values in the `iris` data frame.

**Solution**

```{r}
stripchart(
    iris$Petal.Width ~ iris$Species,
    main = "Petal Width Boxplot",
    xlab = 'Width',
    method = 'jitter',
    jitter = 0.4,
    pch = 20,
    cex = 1.5,
    las = 1,
    col = c('red', 'yellow', 'blue')
)
```



\newpage

End of problem 6

\newpage




# Problem 7: Likert Scale

The marketing team at WiDgT has conducted a survey to determine consumer interest in widgets.

The survey consists of 5 statements:

1. "I think widgets are important and valuable."

2. "I enjoy watching videos about widgets."

3. "I dislike widgets intensely."

4. "I am a better person because I possess many widgets."

5. "Only the dissolute and wicked dabble in widgets."

For each statement, participants can respond with either "Agree", "No Opinion", or "Disagree".

Responses to items 1, 2, and 4 are converted to a numeric score using this table:

| Response     | Score |
|:-------------|:-----:|
| "Agree"      |   3   |
| "No Opinion" |   2   |
| "Disagree"   |   1   |

Responses to items 3 and 5 are converted to a numeric score using this table:

| Response     | Score |
|:-------------|:-----:|
| "Agree"      |   1   |
| "No Opinion" |   2   |
| "Disagree"   |   3   |


Also, participants are invited to join the WiDgT Facebook group, and their response is scored using this table:

| Response   | Score |
|:-----------|:-----:|
| "Yes"      |   2   |
| "No"       |   1   |

The scores on all items are added together to obtain a final score.

Subjects are then categorized into 3 groups based on an interval table:

| Range                  | Interest Level  |
|:-----------------------|:----------------|
| Final Score < 12       | Low Interest    |
| 12 <= Final Score < 15 | Medium Interest |
| 15 <= Final Score       | High Interest   |


## Part (a): Data Analysis

The data for this survey is contained in the file "Problem 7 Data.csv".

Your job is to produce three summaries of this data:

* A table of the relative proportions of the three groups of interest levels, with the values displayed with 2 decimal places.

* A barplot of the relative proportions of the three groups of interest levels.

* A pie chart of the relative proportions of the three groups of interest levels.

You're on your own for this one! You'll have to decide how to structure this analysis, and there are many different approaches that are all valid.

Break your computation into multiple steps, using the features of R notebooks to document your work.

Choose your variable names carefully, and write your code clearly so the TAs can understand what you're doing.

In the end we just want to see the table of relative proportions, the barplot, and the pie chart.

Good luck!!

**Solution**

```{r}
# read in data 
problem.7.data <- read.csv(
    "../Problem Set 9 Data/Problem 7/Problem 7 Data.csv"
)
```

```{r}
# Create lookup tables for each 
lookup.1.2.4 <- c(
    "Agree" = 3,
    "No Opinion" = 2,
    "Disagree" = 1
)

lookup.3.5 <- c(
    "Agree" = 1,
    "No Opinion" = 2,
    "Disagree" = 3
)

facebook.lookup <- c(
    "Yes" = 2,
    "No" = 1
)

final.scores.lookup <- 0


# Get scores for items 1, 2, and 4
item.1.score <- lookup.1.2.4[problem.7.data$item.1]

item.2.score <- lookup.1.2.4[problem.7.data$item.2]

item.4.score <- lookup.1.2.4[problem.7.data$item.4]

# Get scores for items 3 and 5
item.3.score <- lookup.3.5[problem.7.data$item.3]

item.5.score <- lookup.3.5[problem.7.data$item.5]

# Facebook response scores
facebook.score <- facebook.lookup[problem.7.data$item.6]
```

```{r}
# Table of relative proportions of the three groups of interest levels, with 2 digits
interest.levels.vector <- 
    item.1.score +
    item.2.score + 
    item.3.score +
    item.4.score + 
    item.5.score + 
    facebook.score

cut.interest.levels <- cut(
    interest.levels.vector,
    breaks = c(0, 12, 15, Inf),
    labels = c("Low Interest", "Medium Interest", "High Interest")
)

relative.proportions.table <- proportions(table(cut.interest.levels))
```

```{r}
# barplot of the relative proportions of the three groups of interest levels
barplot(
    relative.proportions.table,
    ylim = c(0, 1),
    main = "Relative Proportions",
    xlab = 'Interest',
    ylab = 'y',
    col = 'lavender',
    las = 1
)
```

```{r}
# pie chart of the relative proportions of the three groups of interest levels
pie(
    relative.proportions.table,
    main = "Pie Chart of Relative Proportions",
    col = c('pink', 'lavender', 'lightgreen'),
    clockwise = TRUE
)
```



